### 2018-04-18 paper.Rmd

The data imported into childes-db is subject to data integrity checks to ensure that our import of the corpora is accurate and preferable over ad-hoc parsers developed by many individual researchers.
<!-- MCF: more testing details would be useful -->
In order to evaluate our success in replicating CLAN parses, we compared unigram counts in our database with those outputted by CLAN, the command-line tool built specifically for analysis of transcripts coded in CHAT.
We used the CLAN commands FREQ and MLU to compare total token counts and mean lengths of utterance for every speaker in every transcript and compared these these values to our own using the Pearson correlation coefficient. 
The results of the comparison were .99 and .98 for the unigram count and MLU data, respectively, indicating reliable parsing.
<!-- SM: what were the results of these comparisons?  -->
<!-- AS: technically this is all we do right now SM: what more can we do for low technical investment-->

@@

Many times a researcher will want to investigate the statistics of corpora (e.g., their size, number of utterances, number of tokens) before choosing a target corpus or set of corpora for a project.
This web app is intended to provide a basic overview of various corpora in the CHILDES ecosystem, facilitating familiarity with the different corpora and thus assisting with the overall analytic end-to-end process.
<!-- note that the graphs don't currently include a grouping factor, so adding multiple corpora generates one line, so you can't actually compare two corpora in the same visualization -->

@@

\noindent Each of these functions can be supplied with arguments that restrict the query to a particular subset of the data (e.g. by collection, by corpus, by speaker role, by target child age, etc.) and for a particular version of the data. For more detailed documentation, see the package repository (http://github.com/langcog/childesr).

@@

From this figure, it seems likely that children hear "blue" more frequently early in development, but the trajectories of the "red" and "green" are less clear.
We also do not have a good sense of the errors of these measurements, are limited to just a few colors at a time before the plot becomes too busy, and cannot combine frequencies across speakers. <!-- SM:"busy" seems a bit informal; MCF: I'm ok with it. :) -->
To perform this analysis in a more compelling and complete way, a user can use the `childesr` interface.

@@

Gender differences in language production are well-documented. For example, @eriksson2012differences used parent-report data from ten languages and found that female children had larger vocabularies on average than male children in nearly every language.
Comparable cross-linguistic analysis of naturalistic production data has not been conducted, however, and these differences are easy to explore using `childesr`.
By pulling data from the `transcript_by_speaker` table, users can access derived linguistic measures.
In this worked example we walk through a sample analysis that probes the gender differences in early lexical diversity.

First, we utilize the `childesr` function call `get_speaker_statistics` to pull data relating to the aforementioned derived measures for children and their transcripts.
Note that we exclusively select the data of children, and exclude their caregivers.

@@

The data contained in CHILDES is populated from a diverse array of studies that adhered to varying conditions and paradigms.
This point is particularly salient in our gender analysis due to potential non-independence issues that result from longitudinal studies.
For this reason, we fit a linear mixed effects model (`mod_data`, shown in the plot below) to remedy these concerns.^[All code and analyses are available at https://github.com/langcog/childes-db-paper] 
The plot below displays the average MTLD scores for various children at different ages, split by gender, with a line corresponding to our mixed effects model.

<!-- AS: is this sufficient for mixed effects model explanation? -->

@@

```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x = age_years, y = measure, 
                 color = target_child_sex)) +
  geom_point(alpha = .3) + 
  geom_line(data = mod_data, 
            aes(x = age_years, y = measure, color = target_child_sex)) +
  facet_wrap(~language) + 
  ggthemes::theme_few() + 
  ggthemes::scale_color_ptol(name = "Sex") + 
  xlim(0,max_age) + 
  ylim(0,15) + 
  xlab("Age (years)") + 
  ylab("MTLD") + 
  theme(legend.position = "bottom")
```

This plot reveals a slight gender difference in linguistic productivity in young children (displaying a moderate female advantage), replicating the results of @eriksson2012differences.
The goal of this analysis was to showcase an example of the the steps involved in using `childesr` for exploring the CHILDES dataset, as well as some of the potential hazards in working with a diverse set of corpora that were collected through various experimental paradigms.

<!-- SM: not terribly clear what the hazards were: sparsity and autocorrelation? AS: mostly the fact that there are longitudinal studies in here-->

>>>>

The data imported into childes-db is subject to data integrity checks to ensure that our import of the corpora is accurate and preferable over ad-hoc parsers developed by many individual researchers.
<!-- MCF: more testing details would be useful -->
In order to evaluate our success in replicating CLAN parses, we compared unigram counts in our database with those outputted by CLAN, the command-line tool built specifically for analysis of transcripts coded in CHAT.
We used the CLAN commands FREQ and MLU to compare total token counts and mean lengths of utterance for every speaker in every transcript and compared these these values to our own using the Pearson correlation coefficient.
The results of the comparison were .99 and .98 for the unigram count and MLU data, respectively, indicating reliable parsing.

```{r}
# plot error between clan counts and our counts for speaker unigrams
freq_data <- read_csv("data/freq_testing.csv")

ggplot(freq_data %>%
         mutate(error = abs(clan - db) / clan),
       aes(error)) +
  geom_density() +
  xlim(0,.25)
```

```{r}
# plot error between clan counts and our counts for mlu-w
mlu_data <- read_csv("data/mlu_testing.csv")

ggplot(freq_data %>%
         mutate(error = abs(clan - db) / clan),
       aes(error)) +
  geom_density() +
  xlim(0,.25)
```

<!-- SM: what were the results of these comparisons?  -->
<!-- AS: technically this is all we do right now SM: what more can we do for low technical investment-->

@@

Many times a researcher will want to investigate the statistics of corpora (e.g., their size, number of utterances, number of tokens) before choosing a target corpus or set of corpora for a project.
This web app is intended to provide a basic overview regarding the scale and temporal extent of various corpora in CHILDES, as well as giving researchers insight into the aggregate characteristics of CHILDES.
For example, examining the aggregate statistics reveals that coverage peaks at around 30 months.
<!-- SM: note that the graphs don't currently include a grouping factor, so adding multiple corpora generates one line, so you can't actually compare two corpora in the same visualization -->

@@

\noindent Each of these functions take arguments that restrict the query to a particular subset of the data (e.g. by collection, by corpus, by speaker role, by target child age, etc.) and for a particular version of the data. For more detailed documentation, see the package repository (http://github.com/langcog/childesr).

@@

From this figure, it seems likely that children hear "blue" more frequently early in development, but the trajectories of "red" and "green" are less clear.
We also do not have a good sense of the errors of these measurements, are limited to just a few colors at a time before the plot becomes too crowded, and cannot combine frequencies across speakers.
To perform this analysis in a more compelling and complete way, a user can use the `childesr` interface.

@@

Gender has long been known to be an important factor for early vocabulary growth, with girls learning more words earlier than boys [@huttenlocher1991].
Parent-report data from ten languages suggest that female children have larger vocabularies on average than male children in nearly every language [@eriksson2012differences].
Comparable cross-linguistic analysis of naturalistic production data has not been conducted, however, and these differences are easy to explore using `childesr`.
By pulling data from the `transcript_by_speaker` table, a user has access to the multitude of derived linguistic measures that are often used to evaluate a child's grammatical development.
In this worked example we walk through a sample analysis that probes gender differences in early lexical diversity.

First, we use the `childesr` function call `get_speaker_statistics` to pull data relating to the aforementioned derived measures for children and their transcripts.
Note that we exclusively select the data of children, and exclude their caregivers.

@@

The data contained in CHILDES is populated from a diverse array of studies refelcting varying circumstances of data collection.
This point is particularly salient in our gender analysis due to potential non-independence issues that may emerge from the inclusion of many transcripts from longitudinal studies.
To rememdy these concerns, we fit a linear mixed effects model with gender * age (treated as a quadratic predictor) as fixed effects, child identity as a random intercept, and age + gender by language as a random slope.^[All code and analyses are available at https://github.com/langcog/childes-db-paper]
<!--  mod <- lmer(measure ~ target_child_sex * poly(age_years, 2) - 1 +
                    (1 | target_child_id) +
                    (target_child_sex + age_years | language),
  data = data) -->
<!-- SM: not sure I got the random effects structure right -->
The plot below displays the average MTLD scores for various children at different ages, split by gender, with a line corresponding to the prediction of our fit mixed effects model.

@@

```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x = age_years, y = measure,
                 color = target_child_sex)) +
  geom_point(alpha = .3) +
  geom_line(data = mod_data,
            aes(x = age_years, y = measure, color = target_child_sex)) +
  facet_wrap(~language) +
  ggthemes::theme_few() +
  ggthemes::scale_color_ptol(name = "Sex") +
  xlim(0,max_age) +
  ylim(0,15) +
  xlab("Age (years)") +
  ylab("MTLD") +
  theme(legend.position = "bottom")
```
This plot reveals a slight gender difference in linguistic productivity in young children, replicating the moderate female advantage found by @eriksson2012differences.
The goal of this analysis was to showcase an example of the the steps involved in using `childesr` for exploring the CHILDES dataset, as well as some of the potential hazards -- sparsity and non-independance-- that emerge in working with a diverse set of corpora, many of which were collected in longitudinal studies.
